---
title: "Analysis"
author: "Xiaotong Liu"
editor: visual
format:
  html:
    toc: true
    embed-resources: true
    self-contained-math: true
---

R preparation

```{r}
#| echo: true

if (!requireNamespace("pacman", quietly = TRUE)) { 
  install.packages("pacman")
} 
library("pacman")
p_load(tidyverse, here, GGally, ppcor, lme4, lmtest, patchwork)

## read-in data 
#if the data cannot be read, try df_noise_conjdisj <- read.csv(here::here("data",
#                                         "part2_estimation_task",
#                                         "noise_conj_disj.csv"))
df_noise_conjdisj <- read.csv(here::here("data_and_analysis", "data",
                                         "part2_estimation_task",
                                         "noise_conj_disj.csv"))
#155 participants 

df_ranking <- read.csv(here::here("data_and_analysis", "data",
                                  "part3_ranking_task", 
                                  "result_ranking.csv")) |> 
  filter(id %in% df_noise_conjdisj$id) #143 participants 

df_OSPAN <- read.csv(here::here("data_and_analysis", "data",
                                "part4_WMC_Gf",
                                "df_part4_OSPAN.csv")) |> 
  filter(id %in% df_ranking$id) #87 participants 

df_SSPAN <- read.csv(here::here("data_and_analysis", "data", 
                                "part4_WMC_Gf", 
                                "df_part4_SSPAN.csv")) |>  
  filter(id %in% df_ranking$id) #87 participants

#df_RSPAN <- read.csv(here::here("data_and_analysis", "data", 
#                                 "part4_WMC_Gf", 
#                                 "df_part4_RSPAN.csv")) |>  
#  filter(id %in% df_ranking$id)  #87 participants 
###we decided to not use the RSPAN scores

df_HMT <- read.csv(here::here("data_and_analysis", "data", 
                              "part4_WMC_Gf", 
                              "df_HMT.csv")) |>  
  dplyr::select(id, HMT_score = no_correct) |>  
  dplyr::filter(id %in% df_ranking$id)

#creating a sub-directory for saving the plots
plots_dir <- file.path(getwd(), "plots")
if (!dir.exists(plots_dir)) dir.create(plots_dir, recursive = TRUE)
```

Data exclusion ：

```{r}
#part 2: excluding 2 participants who always responded 100\% or 90\%. 
exclusion_noise <-read.csv(here::here("data_and_analysis", "data",
                                      "part2_estimation_task",
                                      "exclusion_part2_IDs.csv"
))

df_noise_conjdisj <- df_noise_conjdisj |> 
 filter(! id %in% exclusion_noise$id) #|> #153 participants  
 #filter(d_average >= 0 & d_average <= 0.5)

#part 3: excluding 8 (who translated German to English when performing the experiment) 
# + 8(who failed the comprehension check question) = 16 participants
exclusion_ranking <- read.csv(here::here("data_and_analysis",
                                         "data", "part3_ranking_task", 
                                   "exclusion_part3_IDs.csv"))

df_ranking <- df_ranking |> filter(! id %in% exclusion_ranking$id) |> 
  filter(id %in% df_noise_conjdisj$id) #125 participants

#part 4: excluding 2 participants, one was excluded for translating German to English when performing the experiment, another for starting the experiment mutiple times 
exclusion_WMGf <- read.csv(here::here("data_and_analysis", "data",
                                      "part4_WMC_Gf", 
                                      "exclusion_part4_IDs.csv"))

df_OSPAN <- df_OSPAN |> filter(! id %in% exclusion_WMGf$id) 
df_SSPAN <- df_SSPAN |> filter(! id %in% exclusion_WMGf$id)

WMC_result <- df_OSPAN |> 
  full_join(df_SSPAN, by = c("id", "splength")) #85 participants 

#excluding 9 participants, who did not meet the processing-task accuracy criterion (at least 85%) in both the operation and symmetry span322
WMC_result <- WMC_result  |>  
  group_by(id)  |>  
  fill(OSPAN_processing, SSPAN_processing, .direction = "updown")  |> 
  mutate(OSPAN_pass = ifelse(OSPAN_processing >= 0.85, 1, 0),
         SSPAN_pass = ifelse(SSPAN_processing >= 0.85, 1, 0)) |> 
  filter(OSPAN_pass == 1 & SSPAN_pass == 1 )

WMC_result <- WMC_result |> 
  filter(id %in% df_ranking$id) #69 participants 

df_HMT <- df_HMT |>
  filter(id %in% WMC_result$id) #69 participants

```

Demographic information:

```{r}
data_demo <- read.csv(here::here("data_and_analysis",
                                         "data", 
                                 "part3_ranking_task", 
                                   "data_demo.csv"))

data_demo_RQ1 <- data_demo |> 
  filter(id %in% df_ranking$id)

data_demo_RQ1 |> group_by(gender) |> count() 

data_demo_RQ1 |> 
  summarise(
    mean_age = mean(age),
    sd_age = sd(age),
    min_age = min(age),
    max_age = max(age)
  )

data_demo_RQ2 <- data_demo |> 
  filter(id %in% df_HMT$id) 

data_demo_RQ2 |> group_by(gender) |> count()

data_demo_RQ2 |> 
 summarise(
    mean_age = mean(age),
    sd_age = sd(age),
    min_age = min(age),
    max_age = max(age)
  )
```

**Robustness check of the results:**

To assess the robustness of the results, in addition to the analyses reported in the main text, we repeated the analyses under alternative exclusion criteria: We (1) repeated the analyses without excluding any participant in Part 2 (the estimation task) (by skipping the code `df_noise_conjdisj <- df_noise_conjdisj |> filter(! id %in% exclusion_noise$id)`), and (2) repeated the analyses further excluding those participants whose estimated noise parameters lie outside of the theoretical ranges ( by running the code `df_noise_conjdisj <- df_noise_conjdisj |> filter(! id %in% exclusion_noise$id) |> filter(d_average >= 0 & d_average <= 0.5` instead). Across these specifications, the results were similar to the results reported in the main text and our substantive conclusions remained unchanged.

## Research Question 1

Association between the estiamted noise, number of illogical rankings, and additionally the count of conjunction and disjunction fallacies

**Visualization**:

```{r}
# prepare the dataset for the plot 
df_ranking_aggregated <- df_ranking |>  
  group_by(id) |>
  mutate(sum_illogical = sum(logical_pass))  |>  #the number of trials where the responses were illogical
  dplyr::select(id, sum_illogical, get_logic) |>  
  unique()

aggregated_noise_illogical <- df_ranking_aggregated |> 
  left_join(df_noise_conjdisj, by = "id") |>
  drop_na() |> 
  ungroup() |> 
  dplyr::select(id,
         d_average,  
         #number_conj,
         #number_disj,
         sum_illogical,
         number_conj_disj,
         get_logic,
         z1_d,
         z2_d,
         z3_d,
         z4_d,
         z5_d,
         z6_d,
         ) #125 participants 

#descriptive statistics 
( summary_table <- aggregated_noise_illogical |>
  summarise(across(c(d_average, sum_illogical, number_conj_disj, get_logic), 
            list(
              mean = ~mean(.x, na.rm = TRUE),
              sd   = ~sd(.x, na.rm = TRUE),
              min  = ~min(.x, na.rm = TRUE),
              max  = ~max(.x, na.rm = TRUE)
            ), 
            .names = "{.col}_{.fn}")) |>
  mutate(across(everything(), ~round(.x, 2))))

#illogical rankings grouped by event-set condition 
df_ranking_by_condition <- df_ranking |>
  group_by(id, condition) |>
  summarise(number_logical = sum(logical_pass), .groups = "drop")

df_ranking_by_condition |>
  group_by(condition) |>
  summarise(mean_no_logical = mean(number_logical),
            sd_no_logical   = sd(number_logical))
```

```{r}
#| label: ggpairs-noise-ranking-conjdisj
#| fig-width: 6.4
#| fig-height: 6.4
#| fig-cap: "Pairwise plots for the estimated noise parameter, the number of illogical rankings, and the count of conjunction and disjunction fallacies (data from Parts 2 and 3) "

#plot
custom_labels <- c(
  "atop('Estimated Noise', 'Parameter ' * hat(italic(d)))",
  "atop('Frequency of', 'Illogical Rankings')", 
  "atop('Count of Conjunction', 'and Disjunction Fallacies')"
)


lower_lm <- function(data, mapping,
                     se = TRUE, level = 0.95, linewidth = 1,
                     fullrange = TRUE) {
  ggplot(data = data, mapping = mapping) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(method = "lm",
                se = se, level = level,
                linewidth = linewidth, fullrange = fullrange,
                color = "#2297E6" ) 
}

( p1 <- ggpairs(
  aggregated_noise_illogical,
  columns = 2:4,
  columnLabels = custom_labels,
  labeller = "label_parsed",
   upper = list(continuous = wrap("cor", size = 5, digits = 2)),
  lower = list(continuous = wrap(lower_lm,
                                 se = TRUE, level = 0.95, linewidth = 1)),
  diag  = list(continuous = "densityDiag")
) +
  theme_bw(base_size = 14) ) 

#save the plot
#ggsave(
#  filename = file.path(plots_dir, "noise_ranking_conjdisj.pdf"),
#  plot = p1,
#  width = 6.4, height = 6.4,
#  dpi = 300
#)
```

Statistical modelling:

**Pearson correlation** between the estimate noise and the number of illogical rankings:

```{r}
cor.test(aggregated_noise_illogical$d_average, 
         aggregated_noise_illogical$sum_illogical,
                method = "pearson",         
                conf.level = 0.95)
```

**Pearson correlation** between the estimated noise and the count of disjunction and conjunction fallacies:

```{r}
cor.test(aggregated_noise_illogical$d_average, 
aggregated_noise_illogical$number_conj_disj,
                method = "pearson",         
                conf.level = 0.95)
```

**Pearson correlation** between the number of illogical rankings and the count of disjunction and conjunction fallacies:

```{r}
cor.test(aggregated_noise_illogical$sum_illogical, 
aggregated_noise_illogical$number_conj_disj,
                method = "pearson",         
                conf.level = 0.95)

```

**Partial correlation** between the estimated noise parameter and the number of illogical rankings, controlling for the count of conjunction and disjunction fallacies:

```{r}
pc <- pcor.test(aggregated_noise_illogical[,"d_average"], aggregated_noise_illogical[,"sum_illogical"], aggregated_noise_illogical[,"number_conj_disj"], method = "pearson")
pc
```

95% confidence interval:

```{r}
r  <- as.numeric(pc$estimate)
n  <- pc$n
k  <- pc$gp                    # number of control vars
se <- 1 / sqrt(n - k - 3)      # Fisher z SE
z  <- atanh(r)
zcrit <- qnorm(0.975)
ci <- tanh(c(z - zcrit*se, z + zcrit*se))
ci
```

**Partial correlation** between the number of illogical rankings and the count of conjunction and disjunction fallacies, controlling for the estimated noise parameter

```{r}
pc2 <- pcor.test(aggregated_noise_illogical[,"number_conj_disj"], aggregated_noise_illogical[,"sum_illogical"], 
                 aggregated_noise_illogical[,"d_average"], method = "pearson")
pc2
```

95% confidence interval:

```{r}
r2  <- as.numeric(pc2$estimate)
n2  <- pc2$n
k2  <- pc2$gp                    # number of control vars
se2 <- 1 / sqrt(n2 - k2 - 3)      # Fisher z SE
z2  <- atanh(r2)
zcrit2 <- qnorm(0.975)
ci2 <- tanh(c(z2 - zcrit2*se2, z2 + zcrit2*se2))
ci2

```

**GLMM (logit link function; family: binomial) predicting the probability of logical rankings** with predictors: event-set type, the estimated noise parameter (the interaction between the event-set type and the estimated noise parameter, the trial number, as well as the logic rule question)

```{r}
#prepare the dataset for GLMMs
noise_illogical_by_trials <- df_ranking  |>  
  left_join(df_noise_conjdisj, by = "id")  |> 
  dplyr::select(id, f00, condition, logical_pass, 
                d_average, number_conj_disj, get_logic) #125 participants 

```

**Exploratory analysis: Is *even-set condition* a moderator?**

predictors: noise, event-set type, and their interactions

the interaction not significant

```{r}
options(contrasts = c("contr.treatment", "contr.poly"))

glmm_condition_noise_interaction <- glmer(logical_pass ~ condition + d_average + 
                                            condition*d_average +
                                           (1 + condition|id),
                     data = noise_illogical_by_trials,
                     family = binomial(link = "logit"))

summary(glmm_condition_noise_interaction)

get_or_table <- function(model, ci_level = 0.95) {
  # Get coefficients and SEs
  summ <- summary(model)$coefficients
  coefs <- summ[, "Estimate"]
  ses <- summ[, "Std. Error"]
  
  # Calculate critical value based on user-defined CI level
  tail_area <- (1 - ci_level) / 2
  zval <- abs(qnorm(tail_area))
  
  # The Math
  lower <- exp(coefs - zval * ses)
  upper <- exp(coefs + zval * ses)
  or <- exp(coefs)
  
  # Combine and round
  res <- round(cbind(OR = or, Lower = lower, Upper = upper), 3)
  return(as.data.frame(res))
}

get_or_table(glmm_condition_noise_interaction)
```

**Exploratory analysis: trial number as predictor**

predictors: noise, event-set type, f00(number of trials), if participants answered logic question correct or not (get_logical = 1, if the participant answered the question correctly)

```{r}
options(contrasts = c("contr.treatment", "contr.poly"))

glmm_condition_noise_f00 <- glmer(logical_pass ~ condition + d_average + 
                                             f00 +
                                           (1 + condition |id),
                     data = noise_illogical_by_trials,
                     family = binomial(link = "logit"))

summary(glmm_condition_noise_f00)

get_or_table(glmm_condition_noise_f00)
```

**Exploratory analysis:** if the estimated noise parameter $\hat{d}$ explained additional variance in the frequency of illogical rankings beyond the effect of the event-set conditions

predictors: event-set type

```{r}

options(contrasts = c("contr.treatment", "contr.poly"))

glmm_condition <- glmer(logical_pass ~ condition + 
                                           (1 + condition|id),
                     data = noise_illogical_by_trials,
                     family = binomial(link = "logit"))

```

predictors: noise, event-set type

```{r}
options(contrasts = c("contr.treatment", "contr.poly"))

glmm_condition_noise <- glmer(logical_pass ~ condition + d_average + 
                                           (1 + condition|id),
                     data = noise_illogical_by_trials,
                     family = binomial(link = "logit"))

summary(glmm_condition_noise)

get_or_table(glmm_condition_noise)
```

likelihood ratio test:

```{r}
lrtest(glmm_condition, glmm_condition_noise)
```

**Exploratory analysis**: double conjunction fallacy

```{r}
n_total <- nrow(noise_illogical_WMC_Gf_aggregated)

desc <- noise_illogical_WMC_Gf_aggregated |>
  summarise(
    n            = n(),
    mean_count   = mean(number_double_conj, na.rm = TRUE),
    sd_count     = sd(number_double_conj, na.rm = TRUE),
    median_count = median(number_double_conj, na.rm = TRUE),
    min_count    = min(number_double_conj, na.rm = TRUE),
    max_count    = max(number_double_conj, na.rm = TRUE),
    n_any        = sum(number_double_conj > 0, na.rm = TRUE),   # made >= 1 fallacy
    prop_any     = mean(number_double_conj > 0, na.rm = TRUE)   # proportion
  )
print(desc)


test1 <- binom.test(n_any, n_total, p = 0.5, alternative = "greater")
print(test1)

```

## Research Question 2

Can noise predict illogical rankings independently of working memory capacity (WMC) and fluid intelligence (Gf)?

**Visualization:**

```{r}
#prepare the dataset for the plot 
WMC_HMT <- WMC_result |> 
  group_by(id) |> 
  mutate(OSPAN_total = sum(OSPAN_partial_credit, na.rm = TRUE),
         SSPAN_total = sum(SSPAN_partial_credit, na.rm = TRUE)) |> 
  dplyr::select(id, OSPAN_total, SSPAN_total) |> 
  unique() |> 
  mutate(WMC_composite = OSPAN_total + SSPAN_total) |> 
  left_join(df_HMT, by = "id")

noise_illogical_WMC_Gf_aggregated <- aggregated_noise_illogical  |>  
  left_join(WMC_HMT , by = "id")  |> 
  ungroup() |> 
  dplyr::select(id, d_average, WMC_composite, HMT_score, sum_illogical, 
                number_conj_disj, get_logic, z1_d, z2_d, z3_d, z4_d, z5_d, z6_d) |> 
  drop_na() #69 participants 

( summary_table2 <- noise_illogical_WMC_Gf_aggregated |>
  summarise(across(c(d_average, WMC_composite, HMT_score, 
                     sum_illogical, number_conj_disj), 
            list(
              mean = ~mean(.x, na.rm = TRUE),
              sd   = ~sd(.x, na.rm = TRUE),
              min  = ~min(.x, na.rm = TRUE),
              max  = ~max(.x, na.rm = TRUE)
            ), 
            .names = "{.col}_{.fn}")) |>
  mutate(across(everything(), ~round(.x, 2))) )
```

```{r}
#| label: ggpairs-noise-ranking-WMC-Gf
#| fig-width: 7
#| fig-height: 7
#| fig-cap: "Pairwise plots for the estimated noise parameter, WMC score, Gf score and the number of illogical rankings (data from Parts 2, 3 and 4) "

#plot 
custom_labels_noise_illogical_WMC <- c(
  "atop('Estimated Noise', 'Parameter ' * hat(italic(d)))",
   "WMC",
   "Gf",
  "atop('Frequency of', 'Illogical Rankings')"
  #"atop('Count of Conjunction', 'and Disjunction Errors')"
 )



( p2 <- ggpairs(
  noise_illogical_WMC_Gf_aggregated, columns = 2:5,
  columnLabels = custom_labels_noise_illogical_WMC,
  labeller = "label_parsed",
  upper = list(continuous = wrap("cor", size = 4, digits = 2)),
  lower = list(continuous = wrap(lower_lm, se = TRUE, level = 0.95, linewidth = 1)),
  diag  = list(continuous = "densityDiag")
) + 
  theme_bw(base_size = 14) )

#save the plot
#ggsave(
#  filename = file.path(plots_dir, "noise_illogical_HMT_WMC.pdf"),
#  plot = p2,
#  width = 7, height = 7,
#  dpi=300
#)
```

**Pearson correlation** between the number of illogical rankings and Gf score(HMT_score; the score for the HMT task)

```{r}
cor.test(noise_illogical_WMC_Gf_aggregated$HMT_score, noise_illogical_WMC_Gf_aggregated$sum_illogical,
                method = "pearson",          
                conf.level = 0.95)
```

**Pearson correlation** between the number of illogical rankings and WMC score

```{r}
cor.test(noise_illogical_WMC_Gf_aggregated$WMC_composite, noise_illogical_WMC_Gf_aggregated$sum_illogical,
         method = "pearson",
         conf.level = 0.95)
```

**Pearson correlation** between the estimated noise parameter and WMC score

```{r}
cor.test(noise_illogical_WMC_Gf_aggregated$d_average, noise_illogical_WMC_Gf_aggregated$WMC_composite,
         method = "pearson",
         conf.level = 0.95)
```

**Pearson correlation** between the estimated noise parameter and Gf score

```{r}
cor.test(noise_illogical_WMC_Gf_aggregated$d_average, noise_illogical_WMC_Gf_aggregated$HMT_score,
         method = "pearson",
         conf.level = 0.95)
```

**GLMM (logit link function; family: binomial) predicting the probability of logical rankings** with predictors: event-set type, the estimated noise parameter, WMC score and Gf score.

```{r}
#prepare the dataset for GLMMs
noise_illogical_HMT_WMC_trials <- df_ranking |>  
  dplyr::select(id, f00, condition, logical_pass, get_logic) |> 
  left_join(WMC_HMT, by = "id") |> 
  left_join(df_noise_conjdisj, by = "id") |>
  dplyr::select(id, f00, condition, logical_pass, get_logic, WMC_composite, HMT_score, d_average) |> 
    drop_na() #69 participants 

```

predictors: event-set type (condition), noise (d_average), WMC score (WMC_composite), Gf score (HMT_score)

```{r}
glmm_condtion_noise_WMC_Gf <- glmer(logical_pass ~ condition + d_average + 
                         WMC_composite + HMT_score +
                       (1 + condition |id),
                     data = noise_illogical_HMT_WMC_trials,
                     family = binomial(link = "logit"),
                     control = glmerControl(optimizer="bobyqa",
                                   optCtrl=list(maxfun=1e5)))
summary(glmm_condtion_noise_WMC_Gf)
```

Predictors: event-set type (condition), noise (d_average), WMC score (WMC_composite), Gf score (HMT_score), interaction between condition and WMC score, interaction between condition and Gf score

```{r}
glmm_condtion_noise_WMC_Gf_interactions <- glmer(logical_pass ~ condition + d_average + 
                         WMC_composite + HMT_score + 
                         condition*d_average + condition*HMT_score + condition*WMC_composite +
                        (1 + condition |id),
                     data = noise_illogical_HMT_WMC_trials,
                     family = binomial(link = "logit"),
                     control = glmerControl(optimizer="bobyqa",
                                   optCtrl=list(maxfun=1e5)))

summary(glmm_condtion_noise_WMC_Gf_interactions)
```

**Z-transform continuous predictors for the ease of interpretation:**

```{r}
#prepare the dataset:
WMC_mean <- mean(WMC_HMT$WMC_composite) #48.33333
WMC_sd <- sd(WMC_HMT$WMC_composite) #11.36774
hmt_mean <- mean(WMC_HMT$HMT_score) #4.710145
hmt_sd <- sd(WMC_HMT$HMT_score) #1.28454

WMC_HMT_z <- WMC_HMT |> 
  mutate(WMC_z =  (WMC_composite - WMC_mean) / WMC_sd,
         HMT_z = (HMT_score - hmt_mean) / hmt_sd)

df_noise_conjdisj_z <- df_noise_conjdisj |>  
  filter(id %in% WMC_HMT_z$id)

noise_mean <- mean(df_noise_conjdisj_z$d_average)
noise_sd <- sd(df_noise_conjdisj_z$d_average)

df_noise_conjdisj_z <- df_noise_conjdisj_z |>  
  mutate(noise_z = (d_average - noise_mean)/noise_sd) 

noise_illogical_HMT_WMC_trials_z <- df_ranking |> 
  left_join(WMC_HMT_z , by = "id") |> 
  left_join(df_noise_conjdisj_z, by = "id") |> 
  dplyr::select(id, f00, condition, logical_pass, number_conj_disj, 
         noise_z, WMC_z , HMT_z, get_logic) |> 
  drop_na()

#GLMM:
glmm_z_condition_noise_WMC_Gf <- glmer(logical_pass ~ condition + noise_z + 
                         WMC_z + HMT_z +  
                       (1 + condition |id),
                     data = noise_illogical_HMT_WMC_trials_z,
                     family = binomial(link = "logit"),
                     control = glmerControl(optimizer="bobyqa",
                                   optCtrl=list(maxfun=1e5)))
summary(glmm_z_condition_noise_WMC_Gf)
```

Predictor coefficients:

```{r}
round(coef(summary(glmm_z_condition_noise_WMC_Gf)),2)

```

Odds ratios (OR) of predictors:

```{r}
get_or_table(glmm_z_condition_noise_WMC_Gf)
```

**Exploratory analysis**: Did Gf scores differ significantly between participants who correctly identified the logic of the mathematical notations and those who did not?

```{r}

noise_illogical_WMC_Gf_aggregated |> 
  filter(get_logic == "FALSE") |> 
  summarise(mean_Gf = mean(HMT_score),
            sd_Gf = sd(HMT_score))

noise_illogical_WMC_Gf_aggregated |> 
  filter(get_logic == "TRUE") |> 
  summarise(mean_Gf = mean(HMT_score),
            sd_Gf = sd(HMT_score))


t.test(HMT_score ~ get_logic, data = noise_illogical_WMC_Gf_aggregated)

```

**Exploratory analysis**: Correlate each probabilistic expression with WMC and Gf scores

Correlating each probabilistic expression with WMC：

```{r}
r_label <- function(x, y, italic = TRUE) {
  test  <- cor.test(x, y)
  r     <- unname(test$estimate)
  p     <- test$p.value
  stars <- if (p < .001) "***" else
    if (p < .01) "**" else
      if (p < .05) "*" else
        if (p < .10) "." else ""
  
  r_txt <- sub("^(-?)0\\.", "\\1.", sprintf("%.2f", r))
  
  if (italic) paste0("italic(r)~'= ", r_txt, stars, "'")
  else        paste0("r = ", r_txt, stars)
}

wmc_z <- function(var, data, xvar = "WMC_composite",
                  xlab = "WMC Score") {
  
  lab <- r_label(data[[xvar]], data[[var]])
  zname <- toupper(sub("_d$", "", var))
  
  ggplot(data, aes(x = .data[[xvar]], y = .data[[var]])) +
    geom_point(alpha = 0.6, size = 1) +
    geom_smooth(method = "lm", se = TRUE,
                color = "#2297E6", fill = "#2297E6", alpha = 0.1) +
    annotate("text",
             x = -Inf, y = 0.7, hjust = -0.2, vjust = 1.5,
             label = r_label(data[[xvar]], data[[var]]),
             parse = TRUE,
             size = 3.5, color = "#D55E00", fontface = "bold") +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
    labs(x = xlab,
         y = paste(zname, "Value"))+
    theme_bw(base_size = 10) +
    theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"),
          axis.text = element_text(size = 9))
}

plots_wmc <- lapply(c("z1_d", "z2_d", "z3_d", 
                  "z4_d", "z5_d", "z6_d"), wmc_z, 
                data = noise_illogical_WMC_Gf_aggregated)

names(plots_wmc) <- c("z1_d", "z2_d", "z3_d", 
                  "z4_d", "z5_d", "z6_d")

p_wmc <- wrap_plots(plots_wmc[paste0("z", 1:6, "_d")], ncol = 3)
```

Correlating each probabilistic expression with Gf Score：

```{r}
Gf_z <- function(var, data, xvar = "HMT_score",
                  xlab = "Gf Score") {
  
  lab <- r_label(data[[xvar]], data[[var]])
  zname <- toupper(sub("_d$", "", var))
  
  ggplot(data, aes(x = .data[[xvar]], y = .data[[var]])) +
    geom_point(alpha = 0.6, size = 1) +
    geom_smooth(method = "lm", se = TRUE,
                color = "#2297E6", fill = "#2297E6", alpha = 0.1) +
    annotate("text",
             x = -Inf, y = 0.7, hjust = -0.2, vjust = 1.5,
             label = r_label(data[[xvar]], data[[var]]),
             parse = TRUE,
             size = 3.5, color = "#D55E00", fontface = "bold") +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
    labs(x = xlab,
         y = paste(zname, "Value"))+
    theme_bw(base_size = 10) +
    theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"),
          axis.text = element_text(size = 9))
}

plots_Gf <- lapply(c("z1_d", "z2_d", "z3_d", 
                      "z4_d", "z5_d", "z6_d"), Gf_z, 
                    data = noise_illogical_WMC_Gf_aggregated)

names(plots_Gf) <- c("z1_d", "z2_d", "z3_d", 
                      "z4_d", "z5_d", "z6_d")

p_gf  <- wrap_plots(plots_Gf[paste0("z", 1:6, "_d")],  ncol = 3)

p_gf_wmc <- (wrap_elements(p_wmc) / wrap_elements(p_gf)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold", size = 12))

ggsave(
  filename = file.path(plots_dir, "z1_z6_wmc_gf.pdf"),
  plot = p_gf_wmc,
  width = 7, height = 7,
)

```
